---
title: Gaussian Processes
author:  Matt Brigida 
date: 
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
---

Bayesian Machine Learning methods have become very popular recently. The methods lend themselves to updating predictions given new data (which often now some in nearly continuous streams), and they provide distributions of the prediction. In finance, gaussian processes are commonly used in fitting the term structure as well as trading strategies.  [See this paper for background](https://arxiv.org/abs/1903.04841).

We'll first estimate a Gaussian regression using the `GauPro` package in R on a CPU, then using `scikit-learn` in Python on a CPU, and finally estimate the same regression on a GPU in Python using `GPJax` from the `Jax` machine learning framework.

Getting started with Gaussian Processes in easier in R which is why we'll look at an implementation in that language first. See the [GauPro vignettes here](https://cran.r-project.org/package=GauPro).

## Data 

We'll pull crude oil and natural gas data from the EIA's API using my R package updated for EIA's v2 API.

```{r}
library(devtools)
library(GauPro)
library(xts)
```


```{r eval=FALSE, message=FALSE}
install_github("Matt-Brigida/EIAdata", ref = "v2_fix")

library(EIAdata)

key <- source("~/eia_key")$value

weekly_ng <- getEIA("NG.RNGC1.W", key=key)
Sys.sleep(10)
weekly_cl <- getEIA("PET.RCLC1.W", key=key)

data <- merge.xts(weekly_ng, weekly_cl, join="inner")
data
## write to csv to also test in python-------
# write.zoo(data, "./ng_cl_data.csv", sep=",")
```

```{r eval=TRUE, message=FALSE}
data <- read.csv("ng_cl_data.csv")
data <- as.xts(data, order.by=as.Date(data$Index))[, c(2,3)]
```

We'll convert the data into logs and plot:

```{r}
data_post_2000 <- data['2000-01-01/']
ng <- as.numeric(data_post_2000$NG.RNGC1.W)
lng <- log(as.numeric(data_post_2000$NG.RNGC1.W))
cl <- as.numeric(data_post_2000$PET.RCLC1.W)
lcl <- log(as.numeric(data_post_2000$PET.RCLC1.W))

# plot(cl, ng)
plot(lcl, lng)
```

## GauPro in R

<!---
## levels------

gp <- GauPro(as.numeric(data$NG.RNGC1.W), as.numeric(data$PET.RCLC1.W), parallel=FALSE)

gp$predict(5.33)
gp$predict(3.33)
gp$predict(2.33)
gp$predict(1.33)

plot(gp)

## try exponential kernel-----
kern.exp <- Exponential$new(0)
gpk.exp <- GauPro_kernel_model$new(matrix(as.numeric(data$NG.RNGC1.W), ncol=1), as.numeric(data$PET.RCLC1.W), kernel=kern.exp, parallel=FALSE)
plot(gpk.exp)

gpk.exp$predict(2.33)
-->


# Gaussian Regression

```{r}
gpl <- GauPro(lcl, lng, parallel=FALSE)
plot(gpl)
```
We can get a prediction given an input point:

```{r}
gpl$predict(2.)
```

However we also get a standard error for the prediction (`se` below).

```{r}
gpl$predict(2., se=TRUE)
```

```{r}
gpl$predict(3., se=TRUE)
```
```{r}
gpl$predict(4., se=TRUE)
```

```{r}
kern.exp <- Exponential$new(0)
gpk.exp <- GauPro_kernel_model$new(matrix(lcl, ncol=1), lng, kernel=kern.exp, parallel=FALSE)
plot(gpk.exp)
```


```{r}
gpk.exp$predict(2., se=TRUE)
```
```{r}
gpk.exp$predict(3., se=TRUE)
```
```{r}
gpk.exp$predict(4., se=TRUE)
```


## Scikit-Learn in Python

[Here is a good tutorial](https://scikit-learn.org/stable/auto_examples/gaussian_process/plot_compare_gpr_krr.html)

```{python eval=FALSE}
from sklearn.gaussian_process import GaussianProcessRegressor
from sklearn.gaussian_process.kernels import WhiteKernel
import pandas as pd
```

```{python eval=FALSE}
data = pd.read_csv("ng_cl_data.csv")
data
```

```{python eval=FALSE}
kernel = 1.0 * ExpSineSquared(1.0, 5.0) + WhiteKernel(1e-1)
gaussian_process = GaussianProcessRegressor(kernel=kernel)
start_time = time.time()
gaussian_process.fit(training_data, training_noisy_target)
print(
    f"Time for GaussianProcessRegressor fitting: {time.time() - start_time:.3f} seconds"
)

```


## Pyro
