---
title: Gaussian Processes
author: 
date: 
output: html_document
---

```{r}
library(devtools)
library(GauPro)

install_github("Matt-Brigida/EIAdata", ref = "v2_fix")

library(EIAdata)

key <- source("~/eia_key")$value

weekly_ng <- getEIA("NG.RNGC1.W", key=key)
weekly_cl <- getEIA("PET.RCLC1.W", key=key)

data <- merge.xts(weekly_ng, weekly_cl, join="inner")
data
## write to csv to also test in python-------
write.zoo(data, "./ng_cl_data.csv", sep=",")
```

```{r}
ng <- as.numeric(data$NG.RNGC1.W)
lng <- log(as.numeric(data$NG.RNGC1.W))
cl <- as.numeric(data$PET.RCLC1.W)
lcl <- log(as.numeric(data$PET.RCLC1.W))

plot(cl, ng)
# plot(lcl, lng)
```

## see vignettes here: https://cran.r-project.org/package=GauPro

## levels------

gp <- GauPro(as.numeric(data$NG.RNGC1.W), as.numeric(data$PET.RCLC1.W), parallel=FALSE)

gp$predict(5.33)
gp$predict(3.33)
gp$predict(2.33)
gp$predict(1.33)

plot(gp)

## try exponential kernel-----
kern.exp <- Exponential$new(0)
gpk.exp <- GauPro_kernel_model$new(matrix(as.numeric(data$NG.RNGC1.W), ncol=1), as.numeric(data$PET.RCLC1.W), kernel=kern.exp, parallel=FALSE)
plot(gpk.exp)

gpk.exp$predict(2.33)



## logs------

gpl <- GauPro(lcl, lng, parallel=FALSE)
plot(gpl)

gpl$predict(5.33)


kern.exp <- Exponential$new(0)
gpk.exp <- GauPro_kernel_model$new(matrix(lcl, ncol=1), lng, kernel=kern.exp, parallel=FALSE)
plot(gpk.exp)
